---
title: "Artists Profiles and Future Predictions"
author: "Group 13"
format: 
  html:
    code-fold: true
    code-summary: "Show the code"
editor: visual
execute: 
  eval: true
  echo: true
  freeze: true
  warning: false
---

# Loading packages
```{r}
pacman::p_load(tidyverse, jsonlite, SmartEDA, tidygraph, ggraph, igraph, lubridate, janitor, ggthemes, plotly, DT, ggfortify, visNetwork, factoextra, FactoMineR, fmsb, forcats)
```

# 6. Descriptive Statistics

## 6.1. Rising Star Profile

```{r}
# Load pre-processed artist profiles
rising_star_profile <- readRDS("data/processed/artists_profile_complete.rds")

# Render interactive data table
datatable(
  rising_star_profile,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE,
    order = list(list(3, 'desc'))  # Sort by total_works by default
  ),
  caption = htmltools::tags$caption(
    style = 'caption-side: top; text-align: left; font-size: 16px; font-weight: bold;',
    'Table: ', 'Rising Star Profile of Artists'
  )
)
```

# 7. Network Analysis and Visualization
## 7.1. Influence Network for Oceanus Folk Genre

```{r}
# Load pre-processed network visualization data
vis_nodes <- readRDS("data/processed/network_vis_nodes.rds")
vis_edges <- readRDS("data/processed/network_vis_edges.rds")

# Render the interactive network
visNetwork(vis_nodes, vis_edges, height = "700px", width = "100%") %>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visLayout(randomSeed = 42) %>%
  visLegend(useGroups = TRUE, position = "right", main = "Node Types")
```

# 8. Clustering Analysis
## 8.1. Data Preparation

```{r}
# Load artist profiles and filter for clustering
rising_star_profile <- readRDS("data/processed/artists_profile_complete.rds")

cluster_data <- rising_star_profile %>%
  filter(!is.na(first_release)) %>%
  select(name, total_works, notable_works, oceanus_folk_works, 
         first_release, time_to_notability, collaborations) %>%
  drop_na()

cluster_scaled <- cluster_data %>%
  select(-name) %>%
  scale()
```

## 8.2. Determine Optimal Number of Clusters (Elbow Method)

```{r, out.width="100%", out.height="600px"}
set.seed(123)
fviz_nbclust(cluster_scaled, kmeans, method = "wss") +
  labs(title = "Elbow Method for Optimal K")
```

## 8.3. Apply K-Means Clustering

```{r}
set.seed(123)
k_result <- kmeans(cluster_scaled, centers = 4, nstart = 25)

cluster_data <- cluster_data %>%
  mutate(cluster = factor(k_result$cluster))

datatable(cluster_data, options = list(pageLength = 10))
```

## 8.4. Interactive PCA Plot

```{r}
pca_result <- prcomp(cluster_scaled, scale. = TRUE)
pca_data <- as.data.frame(pca_result$x[, 1:2])

pca_data <- pca_data %>%
  mutate(
    cluster = factor(k_result$cluster),
    name = cluster_data$name
  )

pca <- ggplot(pca_data, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(aes(text = paste0(
    "<b>Name:</b> ", name, 
    "<br><b>Cluster:</b> ", cluster
  )), alpha = 0.8, size = 2) +
  stat_ellipse(type = "norm", linetype = 2, aes(group = cluster), show.legend = FALSE) +
  labs(
    title = "Interactive PCA Cluster Plot",
    x = "Principal Component 1",
    y = "Principal Component 2"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.position = "right"
  )

ggplotly(pca, tooltip = "text")
```

## 8.5. Cluster Summary Table

```{r}
cluster_summary <- cluster_data %>%
  group_by(cluster) %>%
  summarise(
    total_works = mean(total_works, na.rm = TRUE),
    notable_works = mean(notable_works, na.rm = TRUE),
    oceanus_folk_works = mean(oceanus_folk_works, na.rm = TRUE),
    first_release = mean(first_release, na.rm = TRUE),
    time_to_notability = mean(time_to_notability, na.rm = TRUE),
    collaborations = mean(collaborations, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

datatable(cluster_summary, options = list(pageLength = 5))
```

# 9. Visualize the careers of three artists
## 9.1. Identify the three Artists

```{r}
# Load pre-processed data
rising_star_profile <- readRDS("data/processed/artists_profile_complete.rds")
nodes_tbl <- readRDS("data/processed/nodes_categorized.rds")

# Select artists
artist_sailor <- rising_star_profile %>%
  filter(str_detect(name, regex("Sailor Shift", ignore_case = TRUE)))

ivy_members <- c("Maya Jensen", "Lila Hartman", "Jade Thompson", "Sophie Ramirez")
artist_peer <- rising_star_profile %>%
  filter(name %in% ivy_members) %>%
  arrange(desc(notable_works)) %>%
  slice(1)

artist_emerging <- rising_star_profile %>%
  filter(!name %in% c(ivy_members, "Sailor Shift")) %>%
  filter(oceanus_folk_works > 0, time_to_notability >= 0) %>%
  arrange(time_to_notability, desc(notable_works)) %>%
  slice(1)

selected_artists <- bind_rows(artist_sailor, artist_peer, artist_emerging)

datatable(selected_artists, options = list(pageLength = 5), 
          caption = "Selected Artists for Career Comparison")
```

## 9.2. Getting their Contributions (Songs/Albums)

```{r}
# Load pre-processed data
edges_tbl_mapped <- readRDS("data/processed/edges_mapped.rds")
id_map <- readRDS("data/processed/id_mapping.rds")

# Prepare artist data
artist_nodes <- selected_artists %>%
  select(artist_id = person_id, artist_name = name)

artist_nodes_indexed <- artist_nodes %>%
  left_join(id_map, by = c("artist_id" = "id")) %>%
  rename(index = index)

relevant_types <- readRDS("data/processed/influence_types.rds")

# Get artist works
artist_edges <- edges_tbl_mapped %>%
  filter(edge_type %in% relevant_types) %>%
  left_join(artist_nodes_indexed, by = c("from" = "index")) %>%
  filter(!is.na(artist_name))

artist_works <- artist_edges %>%
  left_join(
    nodes_tbl %>%
      select(work_id = id, work_name = name, work_type = node_type,
             release_date, genre, notable),
    by = c("to" = "work_id")
  ) %>%
  filter(work_type %in% c("Song", "Album")) %>%
  select(artist_name, work_name, release_date, genre, notable, work_type, edge_type) %>%
  mutate(release_date = as.integer(release_date)) %>%
  arrange(artist_name, release_date)

datatable(
  artist_works,
  rownames = FALSE,
  options = list(
    pageLength = 10,
    autoWidth = TRUE,
    scrollX = TRUE
  ),
  caption = htmltools::tags$caption(
    style = 'caption-side: top; text-align: left; font-size: 16px; font-weight: bold;',
    'Table: ', 'Works Contributed to or Influenced by Selected Artists'
  )
)
```

## 9.3. Plotting Career Timeline

```{r , out.width="100%", out.height="600px"}
top_genres <- artist_works %>%
  count(genre, sort = TRUE) %>%
  slice_head(n = 6) %>%
  pull(genre)

levels_genre_rev <- sort(top_genres, decreasing = TRUE)

artist_subset <- artist_works %>%
  filter(genre %in% top_genres) %>%
  mutate(genre = factor(genre, levels = levels_genre_rev))

ggplot(
  artist_subset,
  aes(x = release_date, y = artist_name, shape = genre, color = notable)
) +
  geom_jitter(width = 0.3, height = 0.3, size = 3, alpha = 0.7) +
  scale_color_manual(values = c(`TRUE` = "#a1d99b", `FALSE` = "#fcbba1")) +
  labs(
    title  = "Timelines (Top Genres Sorted Z–A)",
    shape  = "Genre",
    color  = "Notable"
  ) +
  theme_minimal()
```

## 9.4. Plotting EGO Network Diagram for Top 3 Artists

```{r}
# Load pre-processed graph
graph <- readRDS("data/processed/full_graph.rds")

# Get artist indices
ego_artist_nodes <- nodes_tbl %>%
  filter(node_type == "Person", name %in% selected_artists$name) %>%  
  left_join(id_map, by = c("id" = "id")) %>%
  select(artist_name = name, artist_id = id, index)

ego_indices <- ego_artist_nodes$index

# Extract ego network
ego_subgraph <- graph %>%
  activate(nodes) %>%
  filter(node_is_adjacent(ego_indices) | row_number() %in% ego_indices)

# Prepare visualization data
nodes_data <- ego_subgraph %>%
  activate(nodes) %>%
  as_tibble() %>%
  mutate(
    id = row_number(),
    label = ifelse(!is.na(name), name, node_type),
    group = node_type,
    title = paste0("<b>", label, "</b><br>Type: ", node_type)
  ) %>%
  select(id, label, group, title)

edges_data <- ego_subgraph %>%
  activate(edges) %>%
  as_tibble() %>%
  mutate(
    from = as.integer(from),
    to = as.integer(to),
    label = edge_type,
    title = paste("Edge Type:", edge_type)
  ) %>%
  select(from, to, label, title)

# Plot network
visNetwork(nodes_data, edges_data, height = "600px", width = "100%") %>%
  visOptions(
    highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
    nodesIdSelection = TRUE
  ) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visLayout(randomSeed = 42) %>%
  visLegend(
    useGroups = TRUE,
    position = "right",
    main = "Node Type"
  )
```

# 10. Predicting the next Oceanus Folk stars
## 10.1. Defining Thresholds to Filter Candidates

```{r}
current_year <- 2025

future_stars <- rising_star_profile %>%
  filter(
    oceanus_folk_works >= 5,
    notable_works >= 6,
    !name %in% selected_artists$name,
    first_release >= (current_year - 10),
    time_to_notability <= 3 | is.na(time_to_notability)
  ) %>%
  arrange(desc(notable_works), desc(collaborations)) 

future_stars %>% select(name, total_works, notable_works, oceanus_folk_works, collaborations, time_to_notability)
```

## 10.2. Pick Top 3 Candidates and Justify

```{r}
top_3_future_stars <- future_stars %>% 
  slice_max(order_by = notable_works + collaborations + oceanus_folk_works, n = 3)

top_3_future_stars
```

## 10.3. Radar Chart for Each Candidates

```{r}
radar_data <- top_3_future_stars %>%
  select(total_works, notable_works, oceanus_folk_works, collaborations, time_to_notability) %>%
  as.data.frame()

rownames(radar_data) <- top_3_future_stars$name
colnames(radar_data) <- c("Total", "Notable", "Oceanus\nFolk", "Collabs", "Rapid\nRise")

radar_data[, "Rapid\nRise"] <- max(radar_data[, "Rapid\nRise"], na.rm = TRUE) - 
                               radar_data[, "Rapid\nRise"]

radar_data_scaled <- as.data.frame(lapply(radar_data, function(x) {
  scales::rescale(x, to = c(0, 100))
}))
rownames(radar_data_scaled) <- rownames(radar_data)

par(mfrow = c(1, 3), 
    mar = c(2, 2, 4, 2),
    oma = c(1, 1, 3, 1),
    family = "sans")

for (i in 1:nrow(radar_data_scaled)) {
  artist_name <- rownames(radar_data_scaled)[i]
  artist_values <- radar_data_scaled[i, , drop = FALSE]
  
  radar_individual <- rbind(
    rep(100, ncol(radar_data_scaled)),
    rep(0, ncol(radar_data_scaled)),
    artist_values
  )
  
  radarchart(radar_individual,
             axistype = 1,
             pcol = "darkorange",
             pfcol = rgb(1, 0.5, 0, 0.3),
             plwd = 2,
             cglcol = "gray",
             cglty = 1,
             cglwd = 0.8,
             axislabcol = "darkblue",
             caxislabels = paste0(seq(0, 100, 25), "%"),
             vlcex = 0.85,
             title = artist_name,
             calcex = 0.8,
             cex.main = 1.2)
}
```

